---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---

# Loading required packages

```{r}
# Loading dslabs and tidyverse packages
library("dslabs") 
library("tidyverse")
```

# Loading and checking gapminder data

```{r}
# Looking at help file for gapminder data
help(gapminder)
# Getting an overview of data structure
str(gapminder)
```

```{r}
# Getting summary of data from gapminder
summary(gapminder)
```

```{r}
# Determining the type of gapminder object
class(gapminder)
```

# Processing data

## Creating an object named 'africadata'

```{r}
# Creating an object africadata containing only the African countries
africadata <- gapminder %>% filter(continent == "Africa")

# Looking at the structure and summary of object (africadata)
str(africadata)
summary(africadata)
glimpse(africadata)
```

## Creating an object named 'infantmort_lifeexp'

```{r}
# Creating an object infantmort_lifeexp containing only infant_mortality and life_expectancy variables
infantmort_lifeexp <- africadata %>%
                      select(infant_mortality, life_expectancy)

# Looking at the structure and summary of object (infantmort_lifeexp)
str(infantmort_lifeexp)
summary(infantmort_lifeexp)
glimpse(infantmort_lifeexp)
```

## Creating an object named 'pop_lifeexp'

```{r}
# Creating an object pop_lifeexp containing only population and life_expectancy variables
pop_lifeexp <- africadata %>%
                      select(population, life_expectancy)

# Looking at the structure and summary of object (pop_lifeexp)
str(pop_lifeexp)
summary(pop_lifeexp)
glimpse(pop_lifeexp)
```

# Plotting

```{r}
# Plotting life expectancy as a function of infant mortality
ggplot(data = infantmort_lifeexp,
       (aes(x = infant_mortality, y = life_expectancy))) +
  theme_classic() +
  geom_point(size = 1.0) +
  labs(title = "Life expectancy as a function of infant mortality", 
       x = "Infant Mortality", y = "Life Expectancy (In Years)") +
  theme(axis.text = element_text(face = "bold"), plot.title =   element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 14))
```

```{r}
# Plotting life expectancy as a function of population size
ggplot(data = pop_lifeexp,
       (aes(x = population, y = life_expectancy))) +
  theme_classic() +
  geom_point(size = 1.0) +
  scale_x_continuous(trans = 'log10') +
  labs(title = "Life expectancy as a function of population size", 
       x = "Population (In Logscale)", y = "Life Expectancy (In Years)") +
  theme(axis.text = element_text(face = "bold"), plot.title =   element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 14))
```

# Checking data missingness

```{r}
# Looking at the years having missing data for infant mortality in africadata
infant_mortality_yrs <- africadata %>% 
                        filter(is.na(infant_mortality))
table(infant_mortality_yrs$year)
```

# Data processing using African countries data for the year 2000

```{r}
# Creating a new object named africadata_2000 by using only year 2000 data from africadata object
africadata_2000 <- africadata %>%
                   filter(year == 2000)
# Looking at the structure and summary of object (africadata_2000)
str(africadata_2000)
summary(africadata_2000)
glimpse(africadata_2000)
```

# Plotting using African countries data for the year 2000

## Plot 1

```{r}
# Plotting life expectancy as a function of infant mortality in the year 2000
ggplot(data = africadata_2000,
       (aes(x = infant_mortality, y = life_expectancy))) +
  theme_classic() +
  geom_point(size = 1.0) +
  labs(title = "Life expectancy as a function of infant mortality in the year 2000", 
       x = "Infant Mortality", y = "Life Expectancy (In Years)") +
  theme(axis.text = element_text(face = "bold"), plot.title = element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 14))
```

## Plot 2

```{r}
# Plotting life expectancy as a function of population size in the year 2000
ggplot(data = africadata_2000,
       (aes(x = population, y = life_expectancy))) +
  theme_classic() +
  geom_point(size = 1.0) +
  scale_x_continuous(trans = 'log10') +
  labs(title = "Life expectancy as a function of population size in the year 2000", 
       x = "Population (In Logscale)", y = "Life Expectancy (In Years)") +
  theme(axis.text = element_text(face = "bold"), plot.title =   element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 14))
```

# A simple fit using African countries data for the year 2000

```{r}
# Using lm function to fit linear regression model using life expectancy as the outcome and infant mortality as the predictor for the year 2000
fit1 <- lm(life_expectancy ~ infant_mortality, data = africadata_2000)
summary(fit1)

# Using lm function to fit linear regression model using life expectancy as the outcome and population size as the predictor for the year 2000
fit2 <- lm(life_expectancy ~ population, data = africadata_2000)
summary(fit2)
```

# Conclusion

**Based on the p-values for each fit, we found statistically significant negative correlation between infant mortality and life expectancy in the year 2000 (p-value: 2.83e-08) while we found no statistically significant correlation between population size and life expectancy in the year 2000 (p-value: 0.616).**
