---
title: "MADA Module 4: Data Analysis Exercise"
editor: visual
---

**Source of Data:** The dataset used for this analysis is obtained from the CDC's data website <https://data.cdc.gov/NCHS/Provisional-COVID-19-Deaths-by-Race-and-Hispanic-O/ks3g-spdg>. This dataset contains information about the deaths from COVID-19, pneumonia, influenza, and total deaths from the different states of the US including Puerto Rico during 2020-2023.

**LOADING AND PROCESSING DATA**

```{r}
# Loading the required packages
library(readr)
library(tidyverse)
library(naniar)
library(dplyr)
library(ggplot2)

# Loading the data
covid19mod4 <- read_csv("module4_dataanalysis_exercise/data/raw_data/covid19_deaths_race_hispanic_origin_age.csv")
head(covid19mod4, n = 10)
dim(covid19mod4)
```

**DATA CLEANING**

```{r}
# Looking at the variables
colnames(covid19mod4)

# Changing the variable names to lower case for working conveniance
names(covid19mod4) <- tolower(names(covid19mod4))
colnames(covid19mod4)

# Renaming the variables
covid19mod4 <- covid19mod4 %>%
  rename("age" = "age group",
         "race" = "race and hispanic origin group",
         "covid19_deaths" = "covid-19 deaths",
         "total_deaths" = "total deaths",
         "pneumonia_deaths" = "pneumonia deaths",
         "pneumonia_and_covid19_deaths" = "pneumonia and covid-19 deaths",
         "influenza_deaths" = "influenza deaths",
         "pneumonia_influenza_or_covid19_deaths" = "pneumonia, influenza, or covid-19 deaths")
colnames(covid19mod4)       
```

```{r}
# Let's create an object that contains only the variables that will be used in this data analysis exercise
covid19mod4_main <- covid19mod4 %>% select(state, age, race, covid19_deaths, pneumonia_deaths, influenza_deaths,
                                           pneumonia_and_covid19_deaths, pneumonia_influenza_or_covid19_deaths, total_deaths)
glimpse(covid19mod4_main)

# Checking if there are any missing data  
any(is.na(covid19mod4_main))

# Let's go and remove NAs from each columns of the object `covid19mod4_main`
covid19mod4_main <- covid19mod4_main %>% 
                        drop_na(state, age, race, covid19_deaths, pneumonia_deaths, influenza_deaths,
                               pneumonia_and_covid19_deaths, pneumonia_influenza_or_covid19_deaths, total_deaths)

# Making sure there are no missing values in any columns of the object `covid19mod4_main`
any(is.na(covid19mod4_main))
gg_miss_var(covid19mod4_main)
```

```{r}
# Let's look at the age variable of the object `covid19mod4_main`
unique(covid19mod4_main$age)

# Removing word "years" from `age` column values
covid19mod4_main$age<-gsub(" years","",as.character(covid19mod4_main$age))
unique(covid19mod4_main$age)
class(covid19mod4_main$age)

# Changing '85 and over' age category into '85+' category for our easiness and filtering only the age categories starting `1-4` to `85+` years categories with even 4 years spacing
covid19mod4_main$age[covid19mod4_main$age == '85 and over'] <- '85+'
covid19mod4_main <- 
  filter(covid19mod4_main, age %in% c("1-4", "5-14", "15-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75-84", "85+"))
unique(covid19mod4_main$age)
```

```{r}
# Let's look at the race variable of the object `covid19mod4_main`
unique(covid19mod4_main$race)

# Removing `Non-Hispanic More than one race` and `Unknown` categories of race
covid19mod4_main <- 
  filter(covid19mod4_main, race %in% c("Non-Hispanic White", "Non-Hispanic Black", "Non-Hispanic American Indian or Alaska Native", "Non-Hispanic Asian", "Non-Hispanic Native Hawaiian or Other Pacific Islander", "Hispanic"))
unique(covid19mod4_main$race)
```

```{r}
# Let's look at the state variable of the object `covid19mod4_main`
unique(covid19mod4_main$state)

# Removing `United States`, `Puerto Rico`, `District of Columbia`, and `New York City` so that the object `covid19mod4_main` includes only 50 states data
covid19mod4_main <- covid19mod4_main[-c(which(covid19mod4_main$state == "United States")), ]
covid19mod4_main <- covid19mod4_main[-c(which(covid19mod4_main$state == "Puerto Rico")), ]
covid19mod4_main <- covid19mod4_main[-c(which(covid19mod4_main$state == "District of Columbia")), ]
covid19mod4_main <- covid19mod4_main[-c(which(covid19mod4_main$state == "New York City")), ]
unique(covid19mod4_main$state)

# Since we have no missing data and our data is cleaned as we wanted, the object covid19mod4_main is ready to be used for subsequent data analysis and data visualization.
```

**DATA EXPLORATION AND VISUALIZATION**

```{r}
# Let's look at the number of deaths from COVID-19 as per people's race during 2020-2023
covid19death_race <- covid19mod4_main %>% group_by(race) %>%
 summarize(sum = sum(covid19_deaths, na.rm=T))

# Renaming column `sum` into `covid19_deaths_by_race`
covid19death_race <- covid19death_race %>% rename("covid19_deaths_by_race" = "sum")
covid19death_race <- covid19death_race %>%
  arrange(desc(`covid19_deaths_by_race`))
covid19death_race
```

```{r}
# Plotting the race versus COVID-19 deaths in different states of the US for 2020-23
ggplot(data = covid19death_race, aes(x = reorder(race, -covid19_deaths_by_race), covid19_deaths_by_race,
                                     fill = race)) +
  geom_bar(stat = "identity") +
  geom_col() +
  geom_text(aes(y = covid19_deaths_by_race, label = covid19_deaths_by_race), vjust = -0.5,
            check_overlap = FALSE) +
  labs(title = "COVID-19 Deaths by Race in the US", 
       x = "Race", y = "Number of COVID-19 Deaths") +
  theme(legend.position = "right", 
        plot.title = element_text(hjust=0.5),
        legend.background = element_rect(fill = "white", color = 1),
        axis.text.x = element_blank(),
        axis.title = element_text(size = 14),
        axis.ticks.x = element_blank())
```

**SAVING CLEANED DATASET AS A RDS FILE**

```{r}
# Saving cleaned `covid19mod4_main` as a rds file
saveRDS(covid19mod4_main, "module4_dataanalysis_exercise/data/processed_data/covid19mod4_main.rds")
```

**SUMMARY TABLE**

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
summarizing_table=readRDS("module4_dataanalysis_exercise/data/processed_data/covid19mod4_main.rds")
knitr::kable(summarizing_table)
```

**this section added by Yao Lu**

For finding if there is competition between covid and influenza. First we calculate the correlation coefficient.

```{r}
cor(covid19mod4_main$covid19_deaths,covid19mod4_main$influenza_deaths)
```

Here we find that, there is a high correlation coefficient, but why. We didn't add the total population of the state. So the absolute number of death are confounded by the number of population of state. We can add total population later to our data set, then calculation the relative death rates. Then get correlation coefficient of relative death rate to check the competition.

If we just would like to use this data set. We can still build a model to test if the deaths are related with age. Here I add a new variable called percentage of deaths. Take covid-19 death for example.

```{r}

# Let's look at the number of deaths from COVID-19 as per people's state during 2020-2023
covid19death_state <- covid19mod4_main %>% group_by(state) %>%
 summarize(sum = sum(covid19_deaths, na.rm=T))

#add a covid total column
covid19mod4_main$covidtotal <- NA

table <- table(covid19mod4_main$state)
#calculate the number of rows for each state
norow <-as.data.frame( table[1:50])

#give the value of covidtotal for each state, here we just assume the State are ranked by alphabet, if not we can use order function. 
covid19mod4_main <- covid19mod4_main[order(covid19mod4_main$state),]
covid19death_state <- covid19death_state[order(covid19death_state$state),]
norow <- norow[order(norow$Var1),]

  covid19mod4_main$covidtotal <- rep(covid19death_state$sum[1:50],norow$Freq)

covid19mod4_main$covidperc <- covid19mod4_main$covid19_deaths/covid19mod4_main$covidtotal

```

We will try to model. First, take age as categorical variables. Second, take age as continuous variables.

```{r}
lm1 <- lm(covid19mod4_main$covidperc ~ covid19mod4_main$age)
anova(lm1)

summary(lm1)

TukeyHSD(aov(lm(covid19mod4_main$covidperc ~ covid19mod4_main$age)))
```

In this model, the F test shows that the percentage of covid death are influenced by age group.

Turkey test also shows that there are significant difference between some groups, we can summarize some of the groups with close age to reduce the number of age group.

Then we add a new numerical variable of age. 



```{r}
table(covid19mod4_main$age)

covid19mod4_main <- covid19mod4_main[order(covid19mod4_main$age),]

table(covid19mod4_main$age)

norow1 <- as.data.frame(table(covid19mod4_main$age))

covid19mod4_main$age1 <- rep(c((1+4)/2,(15+24)/2,(25+34)/2,(35+44)/2,(45+54)/2,(5+14)/2,(55+64)/2,(65+74)/2,(75+84)/2, 85),norow1$Freq)


#remove 85+
covid19mod4_main <- 
  filter(covid19mod4_main, !(age1 %in% 85))

#lm

lm2 <- lm(covid19mod4_main$covidperc~covid19mod4_main$age1)

summary(lm2)

```

Here we can see that there is a significant positive trend between age and percentage of covid death.


