---
title: "Fitting"
format: html
editor: visual
---

## FITTING FLU ANALYSIS DATA

**LOADING REQUIRED PACKAGES**

```{r}
library(here)
library(tidyverse)
library(tidymodels)
library(performance)
library(vip)
```

**LOADING DATA**

```{r}
fludata_clean <- readRDS(here("fluanalysis", "data", "fludata_clean.rds"))
glimpse(fludata_clean)
```

**DATA FITTING**

**Fitting a linear model using continuous outcome (Body Temperature)**

```{r}
# Setting up the linear model
lm_model <- linear_reg() %>%
  set_engine("lm")

# Fitting a linear model to the continuous outcome (Body temperature) using only the main predictor of interest (RunnyNose)
lm_fit1 <- lm_model %>% fit(BodyTemp ~ RunnyNose, data = fludata_clean)
lm_fit1

# Looking at the results of lm_fit1
glance(lm_fit1)

# Looking at the additional summary of lm_fit1
tidy(lm_fit1)

# Fitting a linear model to the continuous outcome (Body temperature) using all the predictors
lm_fit2 <- lm_model %>% fit(BodyTemp ~ ., data = fludata_clean)
lm_fit2

# Looking at the results of lm_fit2
glance(lm_fit2)

# Looking at the additional summary of lm_fit2
tidy(lm_fit2)

# Comparing lm_fit1 and lm_fit2 on the basis of model performance
compare_performance(lm_fit1, lm_fit2)
```

**Here, the model with only main predictor of interest (RunnyNose) seem to be a better fit compared to the full model including all predictors as the model with only main predictor of interest has lower AIC of 2329.3 and higher adj. R2 of 0.086.**

**Fitting a logistic model using categorical outcome (Nausea)**

```{r}
# Setting up the logistic model
lm_model1 <- logistic_reg() %>%
  set_engine("glm")

# Fitting a logistic model to the categorical outcome (Nausea) using only the main predictor of interest (RunnyNose)
lm_fit3 <- lm_model1 %>% fit(Nausea ~ RunnyNose, data = fludata_clean)
lm_fit3

# Looking at the results of lm_fit3
glance(lm_fit3)

# Looking at the additional summary of lm_fit3
tidy(lm_fit3)

# Fitting a linear model to the continuous outcome (Body temperature) using all the predictors
lm_fit4 <- lm_model1 %>% fit(Nausea ~ ., data = fludata_clean)
lm_fit4

# Looking at the results of lm_fit4
glance(lm_fit4)

# Looking at the additional summary of lm_fit4
tidy(lm_fit4)

# Comparing lm_fit3 and lm_fit4 on the basis of model performance
compare_performance(lm_fit3, lm_fit4)
```

**Here, the full model including all predictors seem to be a better fit compared to the model with only main predictor of interest (RunnyNose) as the full model has lower AIC of 821.5.**
